opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-11-04 chapters/01_introduction.tex
Updated author + affiliations list

% Manual vs (ML) automated Refinement
Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms. \cite{Ozaki2020} 
Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. Commonly used databases include the "Powder Diffraction File"\cite{GatesRector2019}, the "Inorganic Crystal Structure Database" \cite{Belsky2002} and the "Crystallography Open Database" \cite{Grazulis2009}.
An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
Especially for novel observed structure types, simply matching to a database of known structures will not yield good results. Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement. \cite{Feng2019} 
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals. \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

At the moment, the number of openly available datasets for experimental powder data is very limited.
The main sources for experimental patterns are XXX, YYY, and ZZZ.
Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is x orders of magnitude higher than the largest currently curated experimental dataset and x orders of magnitude larger than the opXRD dataset we present here.

>>>>

% Manual vs (ML) automated Refinement
Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms \cite{Ozaki2020}. 
Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. Commonly used databases include the Powder Diffraction File\cite{GatesRector2019}, the Inorganic Crystal Structure Database \cite{Belsky2002}, and the Crystallography Open Database \cite{Grazulis2009}.
An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
Especially for novel observed structure types, simply matching to a database of known structures will not yield good results. Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement \cite{Feng2019}.
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments \cite{Schopmans2023}[cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

At the moment, the number of openly available datasets for experimental powder data is very limited.
The main sources for experimental patterns are XXX, YYY, and ZZZ.
Further commercial datasets such as KKK and LLL exist, but license fees and restrictions in their use to train publishable models or even publish the training data limit their usability.
Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is likely out of reach.
The most advanced ML models currently are trained on approximately $10^7 - 10^8$ simulated diffractograms, which is - to the best of our knowledge - three orders of magnitude larger than the largest currently curated experimental dataset and even x orders of magnitude larger than the number of unlabeled diffractograms in the opXRD dataset we present here.

### 2024-11-04 chapters/04_our_dataset.tex
Updated author + affiliations list

The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013$\textdegree and an average number of 9190 points measured per pattern. They span a wide chemical space, including organic, inorganic and hybrid structures, including 75 different elements of the periodic table.

>>>>

The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013${\textdegree} and an average number of 9190 points measured per pattern. They span a wide chemical space, including organic, inorganic and hybrid structures, including 75 different elements of the periodic table.

### 2024-11-04 main.tex
Updated author + affiliations list

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Jona Östreicher}
\author[2]{Simon Schweidler}
\author[3]{Adie Alwen}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[2]{Ben Breitung}
\author[4]{Maged Abdelsamie}
\author[4]{Carolin Sutter-Fella}
\author[3]{Andrea Hodge}
\author[1,2]{Pascal Friederich}

@@

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
Kaiserstr. 12, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Kaiserstr. 12, 76131 Karlsruhe, Germany}
\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}

@@

\section{Existing datasets}
\input{chapters/02_existing_datasets}

\section{Machine learning for pXRD}
\input{chapters/03_SOTA_ML}

>>>>

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Jona Östreicher}
\author[2]{Simon Schweidler}
\author[3]{Adie Alwen}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[2]{Ben Breitung}
\author[4]{Maged Abdelsamie}
\author[4]{Gregoire Heymans}
\author[4]{Hanlin Hu}
\author[5]{Alexander Wieczorek}
\author[5]{Siarhei Zhuk}
\author[5]{Sebastiann Siol}
\author[6]{Moritz Wolf}
\author[7]{Arthur Hardiagon}
\author[8]{Bin Cao}
\author[8]{Tong-yi Zhang}
\author[4]{Carolin Sutter-Fella}
\author[3]{Andrea Hodge}
\author[1,2]{Pascal Friederich}

@@

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[3]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
\affil[4]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
\affil[5]{Empa-Swiss Federal Laboratories
for Materials Science and Technology, CH-8600 Dübendorf,
Switzerland}
\affil[6]{Engler-Bunte-Institut \& Institute of Catalysis
Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
\affil[7]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
\affil[8]{Guangzhou Municipal Key Laboratory of Materials Informatics,  Hong Kong University of
Science and Technology (Guangzhou), Guangzhou 511400, Guangdong, China}

@@

\*section{Existing datasets}
\input{chapters/02_existing_datasets}

\*section{Machine learning for pXRD}
\input{chapters/03_SOTA_ML}
